{\bf \huge Wikiplia}:\\ The Free Programming Language\\ That Anyone Can Edit

### 2007-03-16 trunk/wikiplia/paper/wikiplia.tex
syntax

@@

\caption{The definition of the {\sf rate} judgment. The rules for
rating the primitives {\sf string}, {\sf sub}, {\sf substr}, {\sf +},
{\sf -}, {\sf int}, {\sf eq} and {\sf if} are omitted for space. 
%
} \label{fig:xmlrate}
\end{figure*}

Wikiplia is built upon a core calculus of structured data with
primitive revision control. Because we wish to support the freedom to
tinker, the structured data take the form of XML (the {\em extensible}
markup language; Figure~\ref{fig:xml}). Similar to the W3C's XML
Validation,\z\wcite{XML schema} we allow the quality of an XML
document to be assessed via a process called {\em evaluation}, whose
output (if any) is itself an XML document. Selected rules for XML
evaluation are simple and are given in
Figures~\ref{fig:xmleval}~and~\ref{fig:xmlrate}.\z\footnote{XML
documents can be self-correcting through the use of the {\sf handle}
primitive, which detects an invalid document and proceeds along an
alternative path. We omit the rules for this feature, which requires
propagating invalid document status throughout evaluation and thus
complicates the rules substantially.}

Revision control is accessed through the class of {\tt cvs} judgments.
We assume a single global repository, which maps {\em keys} (strings)
to lists of {\em revisions}. A revision is a monotonically increasing
and unique {\em revision number} (integer) paired with an XML
document. The judgment ${\tt cvs commit}\ s\ X = i$ creates a new
revision with revision number $i$ and data $X$ and inserts it in under
the key $s$.

\comment{
  \nocite{alice}

  {\small
  \bibliographystyle{plain}
  \bibliography{wikiplia}
  }
}

>>>>

\newcommand\lb{\ensuremath{[\![}}
\newcommand\rb{\ensuremath{]\!]}}

@@

\caption{Evaluation of XML, part 2. The {\sf rate} judgment assesses a
sequence of XML expressions. The rules for rating the primitives {\sf
string}, {\sf sub}, {\sf substr}, {\sf +}, {\sf -}, {\sf int}, {\sf
eq} and {\sf if} are omitted for space.
%
} \label{fig:xmlrate}
\end{figure*}

Wikiplia is built upon a core calculus of structured data with
primitive revision control. Because we wish to support the freedom to
tinker, the structured data take the form of XML (the {\em extensible}
markup language; Figure~\ref{fig:xml}). Similar to the W3C's XML
Validation,\z\wcite{XML schema} we allow the quality of an XML
document to be assessed via a process called {\em evaluation}, whose
output (if any) is itself an XML document. Selected rules for XML
evaluation (the dynamic semantic markup) are given in
Figures~\ref{fig:xmleval}~and~\ref{fig:xmlrate}.\z\footnote{XML
documents can be self-correcting through the use of the {\sf handle}
primitive, which detects an invalid document and proceeds along an
alternative path. We omit the rules for this feature, which requires
propagating invalid document status throughout evaluation and thus
complicates the rules substantially.}

Revision control is accessed through the class of imperative {\tt cvs}
judgments. We assume a single global repository, which maps {\em keys}
(strings) to lists of {\em revisions}. A revision is a monotonically
increasing and unique {\em revision number} (integer) paired with an
XML document. The judgment ${\tt cvs commit}\ s\ X = i$ creates a new
revision with revision number $i$ and data $X$ and inserts it under
the key $s$. The judgment ${\tt cvs checkout}\ s = X$ fetches the most
recent revision for the key $s$, the document $X$ (if no such key
exists, then the document is invalid). The judgment ${\tt cvs\
checkout\ -\!r}\ i\ s = X$ does the same, but for the specific
revision number $i$ (if no such revision exists, the document is
invalid). Finally, ${\tt cvs\ log}\ s = \vec{X}$ fetches all of the
revision numbers for the key $s$, as a series of integers $\vec{X}$.

\subsection{Syntax}

\begin{figure}[b]
\begin{center}
\begin{tabular}{c}
  \begin{tabular}{rcl}
   {\em E} & ::= & {\tt(} $E_1\ E_2\ \ldots\ E_n$ {\tt)} \\
           & $|$ & {\tt"}string{\tt"} \\
           & $|$ & $n$ \\
           & $|$ & symbol \\
           & $|$ & {\tt'}E \\
  \end{tabular} \\[3em]

  \begin{tabular}{r@{\,\,=\,\,}l}
   \lb {\tt(} $E_1\ \ldots\ E_n$ {\tt)}\rb & $\tag{list}{ \lb E_1 \rb\ \ldots\ \lb E_n \rb}$ \\
   \lb {\tt"}$s${\tt"}\rb & $\tag{string}{s}$ \\
   \lb $n$\rb & $\tag{int}{n}$ \\
   \lb $sym$\rb & $\tag{symbol}{sym}$ \\
   \lb {\tt'}$E$\rb & $\tag{quote}{E}$ \\
  \end{tabular}

\end{tabular}
\end{center}
\caption{The syntax for the XESP syntax. The recursively defined $\lb
\cdot \rb$ operation converts an XESP expression into an XML
document.} \label{fig:xesp}
\end{figure}

XML documents are universally parseable.\z\wcite{Parsing} However,
they are difficult to write and read. Therefore, as usual\wcite{RELAX
NG} we create a new syntax by which humans can write and read programs
and which the computer automatically parses and converts to the easily
parsable XML syntax and back to the new syntax, reducing
complexity.\z\footnote{For efficiency, the Wikiplia implementation is
optimized to lazily perform these translations, so that the document
is never represented in XML form.} This compact syntax is based upon
parentheses rather than tags: The XML document $\tag{list}{X_1\ X_2}$
is instead written ${\tt(}X_1\ X_2{\tt)}$. Note that the closing
parenthesis is not {\em named} as in XML, which makes parsing
difficult because the computer must {\em guess} which parenthesis
belongs to which other parenthesis. We therefore call this compact
syntax XESP because it basically {\em reads the programmer's mind} to
guess what the name of the closing parenthesis should be. The grammar
for XESP is given in Figure~\ref{fig:xesp} along with the translation
to XML documents.
